---
title: "Explore Data Warehouses"
author: "Zixuan Liu"
date: "2023-08-01"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##Question 1
Both fact table and star schema are very inportant for organizing and structuring
data for analyzing.
Fact table could used to store quantitative data which are generally numeric.
Each row could represent a specific event, and each column could represent an
attribute related to the event.The fact table could also includes foreign keys
reference to dimension tables. Each dimension table is connected to the fact table
through a foreign key, and the data will be easier to aggregate and analyze
for multidimensional analysis.
Star schema could be used to design a data warehouse, and it's optimized for OLAP
queries. The fact table is at the center of the schema and surrounded by some
dimension tables. The query will be simplified and become faster.
I think transactional database could be used for OLAP, but generally it might
not be a good choice to use it for OLAP. Transactional database is optimized for
fast update, insertion and so on, and is very useful for OLTP cause OLTP requires
reaction to real-time transactional data. Cause transactional database is not
optimized for complex analytical problems, it may not performing well. Also, the
transactional database is generally normalized and the speed of query might be
slower.

##Qeusion 2
Data warehouse is a centralized repository that stores historical and structured
data for enterprise-wide reporting and analysis. Data mart is the subset of data
warehouse, and it could provide specialized data under specific condition. Data 
lake is a vast and relative flexible storage repository which stores both 
structured and unstructured data. The data structure will be stored in its raw
format, which enables the exploration and analyzaion of diverse data types.
Example of data warehouse: Healthcare organization could use a data warehouse to
store large amount of data including the billing system, health records, patient
feedback and so on. Data could be used to make decisions and imporve the caring
ability. Data warehouse could porvide a comprehensive view of many aspects of the
healthcare organization.
Example of data mart: data mart could provide efficient access to data and 
facilitate the decision-making process. Like the data mart for human resources,
it could contain information like employee information, performence, satisfaction
level and so on. By using this data mart, human resource could make some strategies
and decisions.
Example of data lake: The flexibility of data lake allows organizations process
large amounts of data from diverse source. By analyzing the data of customers,
the customers could be classified into different groups based on their behavior,
preference and so on. This could help target the potential customers, and provide
personalize product.
An article about the differences is shown in the following website https://aws.amazon.com/cn/compare/the-difference-between-a-data-warehouse-data-lake-and-data-mart/
##Question 3
I hope to track under what condition, the flights and birds strike. I also want
to know some details about both the flight, the airport and the birds. To find
out the background of the stike, I need to know the date the strike happened, and 
also under what condition it happened. The condition should contain both the 
weather condition and the flying condition. Also, where did the strike happened
is also important. I also need to include the details of the flights and birds to
make the record of the strike complete.
```{r}
library(knitr)

image_url <- "https://imgur.com/a/sQU5fR2"

knitr::include_graphics(image_url)
```

